### 2025-09-23 paper/sec/authoring-workflow.tex
Progess on workflow.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop \InterpretationAgent workflow (Author intervention states in grey)}
    \label{fig:interpretation-agent}
\end{figure}

The author initially provides the text and accompanying data. The \SuggestionAgent analyses the inputs and
identifies candidate fragments of natural language which could be replaced by Fluid expressions.
Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the \InterpretationAgent.
The entry point is the selection by the author of a fragment of text $s$ to interpret, a fragment which may or
may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
closed loop program synthesis step to check that it is well-formed, and, if so, that it evaluates to the
target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in the original
top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful path through
the loop and returns the system to the ideal state where it is waiting for another top-level interaction from
the author.

@@

\begin{figure}[ht]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}

@@

In the system prompt used to guide the Interpretation Assistant in replacing placeholder tags in user-authored
paragraphs with executable Fluid expressions, we include the following information about the transparent
document under construction, the exact task to perform, and constraints on the expected output. The full
system prompt is given in \appref{system-prompt:interpretation-agent}.

@@

\subsection{Editor Loop}
\label{subsec:editor-loop}
Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
is a \kw{Paragraph} the user is authoring in Fluid. Consists of a sequence of text fragments, some of which
are uninterpreted (plain literals), the remainder have underlying expressions linking the text to raw or derived
data.

Workflow:
\begin{enumerate}
\item User selects (a substring of) one of the literal text fragments, indicating they want to link it in this
way
\item Authoring tool generates expression which is either:
  \begin{enumerate}
  \item computes the selected text (e.g.~``greater than'' might be computed by comparing two numbers)
  \item becomes the formal meaning of the selected text (e.g. the text ``carbon intensity of methane
emissions'' can be understood in a specific context as referring to the numerical value ``34 gCO2eq/kWh''
  \end{enumerate}
\item Some kind of validation step, including both automated validation (e.g. checking for runtime errors) and
user validation; if validation fails, goto (2)
\item Partition code into additional definitions and expression; add definitions to main source file, and
incorporate expression into \kw{Paragraph}
\item Goto (1) with updated editor state
\end{enumerate}

>>>>

The author initially imports the text and accompanying data into the system, which creates a Fluid
representation of the target document. Initially that representation is just a string literal of the form
\kw{"""..."""}, where the triple quotes are the Fluid syntax for an \emph{interpolated string}, i.e.~a string
literal where expressions of the form \kw{\{$e$\}} are permitted within the string. The \SuggestionAgent
analyses the inputs and identifies candidate fragments of natural language which could be replaced by Fluid
expressions. Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the
\InterpretationAgent. The entry point is the selection by the author of a fragment of text $s$ to interpret, a
fragment which may or may not have been highlighted previously by the \SuggestionAgent. The system then
generates a candidate Fluid expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression
is validated through a closed loop program synthesis step to check that it is well-formed, and, if so, that it
evaluates to the target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in
the original top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful
path through the loop and returns the system to the ideal state where it is waiting for another top-level
interaction from the author.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
    \label{fig:interpretation-agent}
\end{figure}

@@

\paragraph{System Prompt.}
For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
supplied paragraph with executable Fluid expressions. It includes the following information about the document
under construction, the exact task to perform, and constraints on the expected output. The full system prompt
is given in \appref{system-prompt:interpretation-agent}.

@@

\rpnote{Now something about the rest of the prompt.}

\begin{figure}[t]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}
